---
title: "EDA Notebook #1"
author: "Ashley E. Mullan"
output: html_notebook
---
**Checklist**  
  1. Explanation of the data structure of the dataset
  2. Three hypotheses about the dataset
  3. Three data visualizations exploring the hypothesis, at least two of which must be multivariate. Each visualization must be in a different format from the other two, and you must have at least one categorical and one continuous visualization.
  4. One clustering example
  5. Conclusions reached for the hypothesis using the data visualizations

It appears that we'll be using Shaw's heat map and cluster, and we will use my diamond and player hit distance.

```{r}
library(tidyverse)
data <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/eda_projects/top_hitters_2019_batted_balls.csv")
data <- mutate(data,
  outcome = as.character(outcome),
  Outcome = case_when(
    outcome == 'field_out' ~ 'Field Out',
    outcome == 'home_run' ~ 'Home Run',
    outcome == 'force_out' ~ 'Force Out',
    outcome == 'single' ~ 'Single',
    outcome == 'triple' ~ 'Triple',
    outcome == 'field_error' ~ 'Field Error',
    outcome == 'double' ~ 'Double',
    outcome == 'sac_fly' ~ 'Sacrifice Fly',
    outcome == 'fielders_choice' ~ 'Fielder\'s Choice',
    outcome == 'grounded_into_double_play' ~ 'Grounded Into Double Play',
    outcome == 'double_play' ~ 'Double Play',
    outcome == 'fielders_choice_out' ~ 'Fielder\'s Choice Out',
    outcome == 'sac_bunt' ~ 'Sacrifice Bunt',
    TRUE ~ outcome
  )
  
)
data <- mutate(data,
  pitch_type = as.character(pitch_type),
  Pitch_Type = case_when(
    pitch_type == 'FF' ~ 'Fastball',
    pitch_type == 'SI' ~ 'Fastball',
    pitch_type == 'FT' ~ 'Fastball',
    pitch_type == 'FC' ~ 'Fastball',
    pitch_type == 'FS' ~ 'Fastball',
    pitch_type == 'SL' ~ 'Offspeed',
    pitch_type == 'CH' ~ 'Offspeed',
    pitch_type == 'CU' ~ 'Offspeed',
    pitch_type == 'KC' ~ 'Offspeed',
    pitch_type == 'KN' ~ 'Offspeed',
    pitch_type == 'EP' ~ 'Offspeed',
    pitch_type == 'NULL' ~ 'NULL',
    TRUE ~ pitch_type
  )
)

data <- mutate(data, 
  player = as.character(player),
  Player = case_when(
    player == "trout" ~ 'Mike Trout',
    player == "yelich" ~ 'Christian Yelich',
    player == "bellinger" ~ 'Cody Bellinger',
    player == "bell" ~ 'Josh Bell',
    player == "gallo" ~ 'Joey Gallo',
    TRUE ~ player
  )
)
head(data)
trout <- data %>% filter(player == "trout")
yelich <- data %>% filter(player == "yelich")
bellinger <- data %>% filter(player == "bellinger")
bell <- data %>% filter(player == "bell")
gallo <- data %>% filter(player == "gallo")
```


```{r}
#Graph Tinkering, some of these are selected in next chunk to tentatively present

#angle and velocity color coded by outcome
data %>% ggplot(aes(x = exit_velocity, y = launch_angle, color = Outcome)) + geom_point()

#hit_x and hit_y color coded by outcome
data %>% ggplot(aes(x = hit_x, y = hit_y, color = Outcome)) + geom_point(size = 1, alpha = 0.4)

#hit_x and hit_y color coded by outcome and faceted by player
data %>% ggplot(aes(x = hit_x, y = hit_y, color = Outcome)) + geom_point(size = 1) + facet_wrap( ~ Player)

#pitch type and batted_ball_type
data %>% ggplot(aes(x = batted_ball_type, fill = Pitch_Type)) + geom_bar(position = "dodge") + facet_wrap( ~ player) + labs(x = "")

#hit distance per pitch type faceted by player
data %>% ggplot(aes(x = hit_distance, color = Pitch_Type)) + geom_histogram(binwidth = 20, position = "dodge") + facet_wrap( ~ Player)

#hit distance by exit velocity color coded by outcome
data %>% ggplot(aes(x = hit_distance, y = exit_velocity, color = Outcome)) + geom_point(size = 0.9) #+ facet_wrap( ~ player)

#launch angle by exit velocity color coded by outcome
data %>% ggplot(aes(x = hit_distance, y = launch_angle, color = Outcome)) + geom_point(size = 0.9)

#heat map of exit velocity and hit distance
data %>% ggplot(aes(x = exit_velocity, y = hit_distance)) + stat_density2d(aes(fill = after_stat(level)), geom = "polygon") + scale_fill_gradient(low = "darkblue", high = "darkorange")

#hit distance by exit velocity color coded by outcome
data %>% ggplot(aes(y = hit_distance, x = exit_velocity, color = Outcome)) + geom_point(size = 0.9, alpha = 0.4) #+ facet_wrap( ~ player)


```

```{r}
#Final Visualization One, pending overlay of diamond
graph1 <- data %>% ggplot(aes(x = hit_x, y = hit_y, color = Outcome)) + 
  geom_segment(x = 0, xend = 45*sqrt(2), y = 0, yend = 45*sqrt(2), size = 2, color = "black") + 
  geom_segment(x = 0, xend = -45*sqrt(2), y = 0, yend = 45*sqrt(2), size = 2, color = "black") + 
  geom_segment(x = 45*sqrt(2), xend = 0, y = 45*sqrt(2), yend = 90*sqrt(2), size = 2, color = "black") +
  geom_segment(x = -45*sqrt(2), xend = 0, y = 45*sqrt(2), yend = 90*sqrt(2), size = 2, color = "black") +
  geom_curve(x = 112, xend = -112, y = 100, yend = 100, curvature = 0.51, color = "black", size = 2) +
  geom_point(size = 1, alpha = 0.6) + 
  labs(title = "Final Hit Position on the Diamond", subtitle = "Color Coded by Play Outcome", x = "X Coordinate of Hit", y = "Y Coordinate of Hit") + 
  theme_bw()
ggsave("Visualization1.png", width = 8, height = 6, path = "C:\\Users\\ashle\\OneDrive\\Documents\\Undergraduate Work\\Summer 2020\\CMSACamp")


#Final Visualization Two
graph2 <- data %>% ggplot(aes(x = hit_distance, fill = Pitch_Type)) + geom_histogram(binwidth = 20, position = "dodge") + facet_wrap( ~ Player) + labs(title = "Hit Distances by Pitch Type", subtitle = "Faceted by Player", x = "Distance of Hit", y = "Count", fill = "Pitch Type") + theme_bw() + theme(legend.position = "bottom")
ggsave("Visualization3.png", width = 8, height = 6, path = "C:\\Users\\ashle\\OneDrive\\Documents\\Undergraduate Work\\Summer 2020\\CMSACamp")

#Final Visualization Three will be a heat/contour map, pending tuning of hyperparameters
#this is a backup, we voted for Shaw's heat map, it fits with our narrative better
data %>% ggplot(aes(x = exit_velocity, y = hit_distance)) + stat_density2d(aes(fill = after_stat(level)), geom = "polygon") + geom_point(alpha = 0.6) + scale_fill_gradient(low = "darkblue", high = "darkorange")
```

