---
title: "EDA Notebook #1"
author: "Ashley E. Mullan"
output: html_notebook
---


```{r}
#library(tidyverse)
data <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/eda_projects/top_hitters_2019_batted_balls.csv")
head(data)
table(data$outcome)
```

```{r}
#angle and velocity color coded by outcome
data %>% ggplot(aes(x = exit_velocity, y = launch_angle, color = outcome)) + geom_point()

#hit_x and hit_y
data %>% ggplot(aes(x = hit_x, y = hit_y, color = outcome)) + geom_point(size = 1)

#pitch type and batted_ball_type
data %>% ggplot(aes(x = batted_ball_type, fill = pitch_type)) + geom_bar(position = "dodge") + facet_wrap( ~ player) + labs(x = "")
```
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
